### 2025-03-11 output/paper.tex
almost working
need to do changes in review_writer

```latex
\documentclass[10pt,journal,compsoc]{IEEEtran}
\usepackage{graphicx}
\usepackage{natbib}
\bibliographystyle{IEEEtranN}

\title{Advancing 3D Reconstruction: The Role of Tri-Hybrid Representation in Enhancing Model Accuracy and Efficiency}
\author{John Doe, \IEEEmembership{Member, IEEE}, Jane Smith, \IEEEmembership{Fellow, OSA}, and Alex Johnson, \IEEEmembership{Member, IEEE}}
\markboth{Journal of Advanced Research in 3D Imaging, Vol. 12, No. 1, January 2024}{Doe \MakeLowercase{\textit{et al.}}: Advancing 3D Reconstruction}

\begin{document}
\maketitle

\begin{abstract}
Recent advancements in 3D reconstruction have leveraged the concept of tri-hybrid representation, integrating geometric, photometric, and semantic information to significantly enhance the accuracy and efficiency of reconstructed models. This paper proposes a comprehensive review of the current methodologies, key innovations, and applications of tri-hybrid representation in 3D reconstruction. We analyze various approaches that combine these three types of data to improve depth perception, object recognition, and scene understanding in complex environments. The review identifies critical gaps in current research, such as the need for more robust integration techniques and the challenge of processing high volumes of data efficiently. Opportunities for future research, including the application of machine learning algorithms and the development of more sophisticated fusion techniques, are discussed. This paper aims to provide a foundational understanding for researchers and practitioners alike, offering insights into the potential of tri-hybrid models to revolutionize 3D imaging technologies.
\end{abstract}

\begin{IEEEkeywords}
3D Reconstruction, Tri-Hybrid Representation, Geometric Data, Photometric Information, Semantic Analysis, Machine Learning.
\end{IEEEkeywords}

\section{Introduction}
\input{sections/introduction.tex}

\section{Literature Review}
\input{sections/literature_review.tex}

\section{Methodology}
\input{sections/methodology.tex}

\section{Results}
\input{sections/results.tex}

\section{Discussion}
\input{sections/discussion.tex}

\section{Conclusion}
\input{sections/conclusion.tex}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\columnwidth]{figure_1}
    \caption{This figure illustrates a comparison of 3D reconstructions generated by different models: AlexNet, VGG, and our proposed Tri-Hybrid model.}
    \label{fig:figure_1}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\columnwidth]{figure_2}
    \caption{This figure presents a comparative visualization of 3D object reconstruction from single input images using different techniques.}
    \label{fig:figure_2}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\columnwidth]{figure_3}
    \caption{This figure illustrates a process for reconstructing 3D point clouds from 2D images. The left column shows the original 2D images, and the right column shows the corresponding 3D point clouds.}
    \label{fig:figure_3}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\columnwidth]{figure_4}
    \caption{This figure illustrates the architecture and process flow of a 3D reconstruction system from 2D images, highlighting the integration of geometric, photometric, and semantic data.}
    \label{fig:figure_4}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\columnwidth]{figure_5}
    \caption{This figure presents a comparative analysis of different methods for reconstructing 3D shapes from images, focusing on the accuracy and computational efficiency.}
    \label{fig:figure_5}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\columnwidth]{figure_6}
    \caption{This figure presents a comparison of different methods for reconstructing 3D models of chairs from 2D images.}
    \label{fig:figure_6}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\columnwidth]{figure_7}
    \caption{The figure illustrates a sophisticated architecture for a diffusion model used in generating 3D representations from single images.}
    \label{fig:figure_7}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\columnwidth]{figure_8}
    \caption{The figure illustrates a system for 3D scene understanding from images. It consists of two main components: object detection and scene reconstruction.}
    \label{fig:figure_8}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\columnwidth]{figure_9}
    \caption{This figure presents an analysis of different methods for reconstructing 3D shape, surface normals, and texture from a single image.}
    \label{fig:figure_9}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\columnwidth]{figure_10}
    \caption{This figure presents a comparison between synthetic data and real-world data in the context of 3D object reconstruction.}
    \label{fig:figure_10}
\end{figure}

\bibliography{references}

\end{document}
```